---
title: "PAID Code"
author: "Alexander Starlinger"
date: "2023-12-17"
output: html_document
---

# 1. Packages & Imports

```{r load packages }

library(readxl)
library(tidyverse)
library(lme4)
library(reshape2)
library(lmerTest)
library(emmeans)
library(r2mlm)
library(psych)
library(lmtest)
library(sjPlot)
library(haven)
library(estimatr)
library(flexplot)
library(flextable)

options(scipen = 999)

```

```{r import data}

MA_data <- read.csv2("data/util_judgment_data_wide.csv")
MA_data_long <- read.csv2("data/util_judgment_data_long.csv")

```

# 2. Descriptives & Reliability

utilitarianism.

```{r util descriptives}

MA_data_long %>%
  group_by(self_cwc, inv_cwc) %>%
  summarize(mean = round(mean(UtilScore), 2),
            sd = round(sd(UtilScore), 2))

```

## 2.1. Sample Description

```{r sample description}

socDem <- MA_data %>%
  select(sex, age, highest_education, student, nationality)

# 1. N
nrow(socDem)

# 2. Age
describe(socDem$age)

# 3. sex
socDem$sex %>% table()/nrow(socDem)

# 4. edu; Note: <= 3: primary; 4: secondary; >= 5: tertiary
eduTable <- socDem$highest_education %>% table()/nrow(socDem)

## primary: 19%; secondary: 46.3%; tertiary: 34.7%

# 5. student
table(socDem$student)/nrow(socDem) 

# 6. nationality
table(socDem$nationality)/nrow(socDem)


```

## 2.2 Descriptives

descriptives UtilScores.

```{r descriptives: UtilScores means}

# means per parcel

MA_data_long %>%
  group_by(inv_cwc, self_cwc) %>%
  summarize(mean = round(mean(UtilScore, na.rm = TRUE), 2),
            sd = round(sd(UtilScore, na.rm = TRUE), 2))

```

Descriptives predictors.

```{r descriptives: reliability}

## item objects
LTS_items <- MA_data %>%
  select(GM02_01:GM02_12)
NPI_items <- MA_data %>%
  select(GN01_01:GN01_15)
HSNS_items <- MA_data %>%
  select(VN02_01:VN02_10)
MACH_items <- MA_data %>%
  select(MA02_01:MA02_03, MA02_04_R, MA02_05)
DC_items <- MA_data %>%
  select(DC01_01_R, DC01_02, DC01_03_R, DC01_04, DC01_05_R, DC01_06,
         DC01_07_R, DC01_08, DC01_09_R, DC01_10, DC01_11, DC01_12, DC01_13_R, DC01_14_R, DC01_15, DC01_16_R)
SRP_items <- MA_data %>%
  select(PS01_01:PS01_03, PS01_04_R, PS01_05:PS01_10, PS01_11_R, PS01_12:PS01_16, PS03_17_R, PS03_18_R, PS03_19_R, PS03_20:PS03_31)

## keys für subscalen

fih <- c(1:4)
hu <- c(5:8)
kant <- c(9:12)

lp <- c(1:3, 5:7, 10:12, 14)
g <- c(4, 8, 9, 13, 15)

stj <- c(1:3, 7, 9)
e <- c(4:6, 8, 10)

ab <- c(22:31)
its <- c(1, 3, 5, 9, 10, 12:14, 20)
im <- c(2, 6, 15, 16:19)
ca <- c(4, 7, 8, 11, 21)

## rel

# scales
lts_total_rel <- round(alpha(LTS_items)$total$raw_alpha, 3)
srp_total_rel <- round(alpha(SRP_items, keys = c("PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R"))$total$raw_alpha, 3)
npi_total_rel <- round(alpha(NPI_items)$total$raw_alpha, 3)
hsns_total_rel <- round(alpha(HSNS_items)$total$raw_alpha, 3)
mach_total_rel <- round(alpha(MACH_items)$total$raw_alpha, 3)
dc_total_rel <- round(alpha(DC_items)$total$raw_alpha, 3)

# subscales

lts_fih_rel <- round(alpha(LTS_items[, fih])$total$raw_alpha, 3)
lts_hu_rel <- round(alpha(LTS_items[, hu])$total$raw_alpha, 3)
lts_kant_rel <- round(alpha(LTS_items[, kant])$total$raw_alpha, 3)
srp_ab_rel <- round(alpha(SRP_items[, ab], keys = c("PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R"))$total$raw_alpha, 3)
srp_its_rel <- round(alpha(SRP_items[, its], keys = c("PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R"))$total$raw_alpha, 3)
srp_im_rel <- round(alpha(SRP_items[, im], keys = c("PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R"))$total$raw_alpha, 3)
srp_ca_rel <- round(alpha(SRP_items[, ca], keys = c("PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R"))$total$raw_alpha, 3)
npi_lp_rel <- round(alpha(NPI_items[, lp], keys = c("PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R"))$total$raw_alpha, 3)
npi_g_rel <- round(alpha(NPI_items[, g], keys = c("PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R", "PS01_04_R"))$total$raw_alpha, 3)
hsns_stj_rel <- round(alpha(HSNS_items[, stj])$total$raw_alpha, 3)
hsns_e_rel <- round(alpha(HSNS_items[, e])$total$raw_alpha, 3)


rel <- c(lts_total_rel, lts_fih_rel, lts_hu_rel, lts_kant_rel, npi_total_rel, npi_lp_rel, npi_g_rel, hsns_total_rel, hsns_stj_rel, hsns_e_rel,
         srp_total_rel, srp_ab_rel, srp_im_rel, srp_its_rel, srp_ca_rel, mach_total_rel, dc_total_rel)

# items rausschmeißen

MA_data <- MA_data %>%
  select(!(c(nationality, highest_education, student, GM02_01:PS03_31, MA02_01:DC01_16_R)))

```

```{r descriptives: Konstrukte}

scales <- MA_data %>%
  select(LTS_total:DC_total)

descriptives <- MA_data %>%
  select(LTS_total:DC_total) %>%
  describe() %>%
  select(c(mean, sd))

## append reliability

descriptives$alpha <- rel
descriptives

```

```{r table 1 primary analysis}

preds  <- MA_data %>% select(LTS_FIH:NPI_total, HSNS_STJ:SRP_total, MACH_total:DC_total)
preds <- preds %>% 
  relocate(., after = c("DC_total", "SRP_total", "MACH_total", "NPI_total", "HSNS_E", "HSNS_STJ", "LTS_FIH", "LTS_H", "LTS_K")) 
colnames(preds) <- c("DC_total", "SRP_total", "MACH_total", "NPI_total", "HSNS_E", "HSNS_STJ", "LTS_FIH", "LTS_H", "LTS_K")


tab1 <- preds %>%
  describe() %>%
  select(c(mean, sd)) %>% round(., 2)

tab1_alpha <- descriptives[colnames(preds), ] %>% select(alpha)
tab1$alpha <- round(tab1_alpha$alpha, 2)

pred_cor <- round(corr.test(preds)$r, 2)
pred_cor_p <- round(corr.test(preds)$p, 5)
pred_cor_p


pred_cor[upper.tri(pred_cor, diag = TRUE)] <- NA

tab1 <- cbind(tab1, pred_cor)

tab1

```

# 3. Primary Analysis

Models:

0) Nullmodel (ICC estimate)
1) Covariates + SI/PI + DC
2) random slopes PI
3) Dark Trait block
4) LTS block
5) CL-I SRP*PI

# 3.1. Models

```{r unconditional model}

null_model <- lmer(UtilScore ~ 1 + (1|ID), data = MA_data_long, REML = FALSE)

summary(null_model) 
tab_model(null_model)

```

1) Covariates + SI*PI + DC

```{r DC model}

DC_model <- lmer(UtilScore ~ sex + age + inv_cwc*self_cwc + DC_total_gmc + (1|ID), data = MA_data_long)

summary(DC_model)
tab_model(DC_model)

```

2) + RE inv_cwc.

```{r DC-RE model}

DC_RE_model <- lmer(UtilScore ~ sex + age + inv_cwc*self_cwc + DC_total_gmc + (1 + inv_cwc|ID), data = MA_data_long)

summary(DC_RE_model)
tab_model(DC_RE_model)

anova(DC_model, DC_RE_model) 

```

3) Dark Traits Block.

```{r DT model}

DT_model <- lmer(UtilScore ~ sex + age + inv_cwc*self_cwc + DC_total_gmc + HSNS_E_gmc + HSNS_STJ_gmc + MACH_total_gmc + SRP_total_gmc + NPI_total_gmc + (1 + inv_cwc|ID), data = MA_data_long)

summary(DT_model)
tab_model(DT_model)

anova(DC_RE_model, DT_model) 


```

4) LTS Block.

```{r LTS model}

LTS_model <- lmer(UtilScore ~ sex + age + inv_cwc*self_cwc + DC_total_gmc + SRP_total_gmc + MACH_total_gmc + NPI_total_gmc + HSNS_E_gmc + HSNS_STJ_gmc + LTS_FIH_gmc + LTS_H_gmc + LTS_K_gmc +  (1 + inv_cwc|ID), data = MA_data_long)

summary(LTS_model)
tab_model(LTS_model)

anova(DT_model, LTS_model) # no improvement

```

5) CL-I SRP*inv_cwc.

```{r full model}


full_model <- lmer(UtilScore ~ sex + age + inv_cwc*self_cwc + DC_total_gmc + SRP_total_gmc + MACH_total_gmc + NPI_total_gmc + HSNS_E_gmc + HSNS_STJ_gmc + LTS_FIH_gmc + LTS_H_gmc + LTS_K_gmc + inv_cwc*SRP_total_gmc +  (1 + inv_cwc|ID), data = MA_data_long)

summary(full_model)
tab_model(full_model)

anova(full_model, LTS_model) 

```

# 3.2 EMMEANS

simple slopes SRP total.

```{r simple slopes}

emtrends(full_model, specs = "inv_cwc", var = "SRP_total_gmc") %>%
  test()

```

```{r EMMeans SI*PI}

emm_SIPI <- emmeans(full_model, ~ inv_cwc*self_cwc)
pairs(emm_SIPI)

# calculate d
eff_size(emm_SIPI, sigma(full_model), edf = Inf)

```

# 3.3 Explained Variance

explained variance: r2mlm. 

Note: r2mlm needs numerical predictors. Refit.

```{r R^2: r2mlm}

MA_data_num <- MA_data_long %>% mutate(sex = as.numeric(sex),
                                       self_cwc = as.numeric(self_cwc),
                                       inv_cwc = as.numeric(inv_cwc))

full_model_num <- lmer(UtilScore ~ sex + age + inv_cwc*self_cwc + DC_total_gmc + SRP_total_gmc + MACH_total_gmc + NPI_total_gmc + HSNS_E_gmc + HSNS_STJ_gmc + LTS_FIH_gmc + LTS_H_gmc + LTS_K_gmc + inv_cwc*SRP_total_gmc +  (1 + inv_cwc|ID), data = MA_data_num)

r2mlm(full_model_num)


```

# 3.4 additional models

MACH as single predictor.

```{r MACH}

MACH_model <- lmer(UtilScore ~ sex + age + inv_cwc*self_cwc + DC_total_gmc + MACH_total_gmc +  (1 + inv_cwc|ID), data = MA_data_long)

summary(MACH_model)
tab_model(MACH_model)

anova(DC_RE_model, MACH_model) 

```

LTS as single predictor.

```{r LTS total}

LTS_total_model <- lmer(UtilScore ~ sex + age + inv_cwc*self_cwc + DC_total_gmc + SRP_total_gmc + MACH_total_gmc + NPI_total_gmc + HSNS_E_gmc + HSNS_STJ_gmc + LTS_total_gmc + (1 + inv_cwc|ID), data = MA_data_long)

summary(LTS_total_model)
tab_model(LTS_total_model)

anova(DT_model, LTS_total_model)

```


# 4. Supplementary Analyses

  1) single level regression für SI*DC
  2) facet-level LMM (SRP-III)
  3) facet-level: F1 vs. F2


## 4.1 Single-Level Regression

```{r single-level model: estimatr}

single_level_mod <- lm_robust(UtilScore ~ sex + age + 
                                inv_cwc*self_cwc + 
                                DC_total_gmc + SRP_total_gmc + MACH_total_gmc + NPI_total_gmc +
                                HSNS_E_gmc + HSNS_STJ_gmc + LTS_FIH_gmc +  LTS_H_gmc + LTS_K_gmc +
                                DC_total_gmc*self_cwc +
                                SRP_total_gmc * inv_cwc,
                                cluster = ID, data = MA_data_long)

summary(single_level_mod)
tab_model(single_level_mod)

```

```{r single-level model: probe interaction}

emtrends(single_level_mod, specs = "self_cwc", var = "DC_total_gmc") %>% 
  test()

```

Check if interaction self-interest*SRP improves model fit.

``` {r single-level full}

single_level_mod_full <- lm_robust(UtilScore ~ sex + age + 
                                inv_cwc*self_cwc + 
                                DC_total_gmc + SRP_total_gmc + MACH_total_gmc + NPI_total_gmc +
                                HSNS_E_gmc + HSNS_STJ_gmc + LTS_FIH_gmc +  LTS_H_gmc + LTS_K_gmc +
                                DC_total_gmc*self_cwc +
                                SRP_total_gmc * inv_cwc +
                                SRP_total_gmc * self_cwc,
                                cluster = ID, data = MA_data_long)

summary(single_level_mod_full)

# LRT
waldtest(single_level_mod, single_level_mod_full)


```

## 4.2 SRP Facet-Level Model

```{r SRP Model}

SRP_model <- lmer(UtilScore ~ sex + age + self_cwc * inv_cwc + SRP_AB_gmc + SRP_ITS_gmc + SRP_IM_gmc + SRP_CA_gmc + (1 + inv_cwc | ID) + inv_cwc * SRP_AB_gmc + inv_cwc * SRP_IM_gmc, data = MA_data_long)

summary(SRP_model)
tab_model(SRP_model)
```

simple slopes AB/IM.

```{r simple slopes SRP-AB}

emtrends(SRP_model, specs = "inv_cwc", var = "SRP_AB_gmc") %>% 
  test()
emtrends(SRP_model, specs = "inv_cwc", var = "SRP_IM_gmc") %>% 
  test()

```

## 4.3. SRP: two-factor (primary vs. secondary psychopathy)

```{r SRP twofactor model}

SRP_twofactor_model <- lmer(UtilScore ~ sex + age + self_cwc * inv_cwc + SRP_F1_gmc + SRP_F2_gmc + (1 + inv_cwc | ID) + inv_cwc * SRP_F1_gmc  + inv_cwc * SRP_F2_gmc, data = MA_data_long)

summary(SRP_twofactor_model)
tab_model(SRP_twofactor_model)

```

# 5. Figures

## 5.1 Figure 1

means.

```{r barplot means}

MA_data_long_rec <- MA_data_long %>%
  mutate(self_interest = factor(self_interest, labels = c("no-individual-benefit", "individual-benefit")),
         involvement = factor(involvement, labels = c("impersonal", "personal")))

descr <- MA_data_long_rec %>%
  group_by(self_interest, involvement) %>%
  summarize(mean = round(mean(UtilScore), 2),
            sd = round(sd(UtilScore), 2),
            n = n(),
            se = sd/sqrt(n))

dodge <- position_dodge(0.9)

ggplot(data = descr, aes(x = self_interest, y = mean, fill = involvement)) +
  geom_bar(stat = "identity", position = dodge, color = "black", size = 0.5) +
  geom_errorbar(aes(ymin = mean - 1.96*se, ymax = mean + 1.96*se), position = dodge, width = 0.3) + 
  ylim(-2, 2) +
  theme_minimal() +
  scale_fill_grey() +
  xlab("self interest") +
  ylab("Utilitarianism (mean)")



```

Cohen's d.

```{r EMMeans für SI*PI}

emm_SIPI <- emmeans(full_model, ~ inv_cwc * self_cwc)

# d
tmp <- eff_size(emm_SIPI, sigma(full_model), edf = Inf)
tmp <- tmp[c(1:2, 5:6)]
eff_sizes <- summary(tmp)

#change sign
eff_sizes  <- eff_sizes %>% 
  mutate(effect.size = effect.size*-1,
         LCL = asymp.UCL*-1,
         UCL = asymp.LCL*-1)

# get plotting info
eff_sizes <- eff_sizes %>% 
              select(effect.size, LCL, UCL) %>%
              mutate(effect = factor(rep(c("involvement", "self-interest"), 2)),
              level = factor(rep(c("impersonal/\nno individual benefit", "personal/\nindividual benefit"), each = 2)))

eff_sizes$level <- relevel(eff_sizes$level, "impersonal/\nno individual benefit")

#plot
dodge <- position_dodge(0.9)

ggplot(data = eff_sizes, aes(x = effect, y = effect.size, fill = level)) + 
  geom_bar(stat = "identity", position = dodge, color = "black", size = 0.5) +
  geom_errorbar(aes(ymin = LCL, ymax = UCL), position = dodge, width = 0.3) +  
  theme_minimal() +
  scale_fill_grey(start = 0.8, end = 0.2) +
  xlab("Dilemma characteristic") +
  ylab("Standardized mean difference (d)")

```

## 5.2. Figure 2

Interaction plots. Recode labels.

```{r interaction plots prep}

MA_data_long_rec <- MA_data_long %>%
  mutate(involvement = factor(involvement, labels = c("impersonal", "personal")),
         self_interest = factor(self_interest, labels = c("no individual benefit", "individual benefit")))

```

a) DC*SI.

```{r interaction plot DC*SI}

ggplot(data = MA_data_long_rec, aes(x = DC_total_gmc, y = UtilScore, color = self_interest)) +
    geom_smooth(method = "lm", se = FALSE) + 
    theme_minimal() + 
    ylim(-2, 2) +
    scale_color_grey(name = "self interest", start = 0.8, end = 0.2) +
    xlab("D (standardized)") +
    ylab("Utilitarianism")

```

b) SRP*PI.

```{r interaction plot SRP*PI}

ggplot(data = MA_data_long_rec, aes(x = SRP_total_gmc, y = UtilScore, color = involvement)) +
    geom_smooth(method = "lm", se = FALSE) + 
    theme_minimal() + 
    ylim(-2, 2) +
    scale_color_grey(start = 0.8, end = 0.2) +
    xlab("Psychopathy (standardized)") +
    ylab("Utilitarianism")

```
